---
title: "Get started: installation to interpretation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started: installation to interpretation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(drfarm)
site_link <- function(label, path) {
  if (identical(Sys.getenv("IN_PKGDOWN"), "true")) {
    paste0("[", label, "](", path, ")")
  } else paste0(label, " (on the documentation site)")
}
```

## Install the matching source

These pages describe the **0.1.0.9002 development candidate**. Public GitHub main
still contains historical 0.1.0. Install the candidate before using its new
diagnostics, weighted coefficient option or Gaussian ECM reference.

On the documentation site, download the
`r site_link("candidate source archive", "../downloads/drfarm_0.1.0.9002.tar.gz")`.
From the directory containing that file, run in R:

```{r installation, eval=FALSE}
install.packages(c("glasso", "glmnet", "psych", "Rcpp"))
install.packages("drfarm_0.1.0.9002.tar.gz", repos = NULL, type = "source")
library(drfarm)
packageVersion("drfarm")
```

For the recorded dependency versions, use the full source checkout instead:

```sh
Rscript --vanilla tools/install-dependencies.R
R CMD INSTALL .
```

R ≥ 4.3.0 and a C++ compilation toolchain are required. The current local checks
use **R 4.3.3 on Ubuntu 24.04**. Windows and macOS are not locally checked in
this sprint. On Windows, use the Rtools matching your R installation; on macOS,
install the Xcode command-line tools. If compilation fails, retain its complete
log and `sessionInfo()`; check that Rcpp, glasso, glmnet and psych load first.

To reproduce the historical source specifically, use the pinned GitHub commit:

```{r historical-install, eval=FALSE}
install.packages("remotes")
remotes::install_github(
  "lapsumchan/drfarm@be6d52ee796161e732f398da5eadfc3d40812f34"
)
```

That source has the original methods, but lacks the candidate's finite-budget
diagnostics, weighted solver and ECM reference. It is not a CRAN release.

## Start with the original bundled data

The original demonstration contains 500 participants, 10 predictors and 5
continuous outcomes. Its documented latent rank is two; we use `k = 2` here.
The original simulation generator and seed are not bundled, so this reproduces
the software example rather than the entire original simulation study.

```{r data}
data("drfarm.dat", package = "drfarm")
X <- drfarm.dat$X
Y <- drfarm.dat$Y
rbind(X = dim(X), Y = dim(Y), generating_Theta = dim(drfarm.dat$Theta.t))
```

The generating `Theta.t` is p × q. The fitted `Theta` is q × p: outcome rows,
predictor columns. Orientation and scale must both match before a comparison.

## Reproduce the original fitting workflow

The original main example uses the default 100-cell remMap grid to initialize
coefficients, estimates predictor precision, and fits the default 25-cell DrFARM
grid. The calls below retain that procedure and its selection rule.

```{r fit}
set.seed(20260909)
initial <- remMap.whole(X, Y)
precision <- precM(X)
fit.warnings <- character()
fit <- withCallingHandlers(
  DrFARM.whole(
    X, Y, initial$Theta0, precision, k = 2,
    lambda1.opt = initial$lambda1.opt,
    lambda2.opt = initial$lambda2.opt
  ),
  warning = function(w) {
    fit.warnings <<- c(fit.warnings, conditionMessage(w))
    invokeRestart("muffleWarning")
  }
)
table(fit.warnings)
```

Repeated warnings are counted above because they are part of the result.
The handler only collects messages; it does not change the fit. For new workflows,
choose a finite `max.iter`; this original-workflow reproduction keeps the defaults.
Inspect the returned status before interpreting the coefficients:

```{r status}
fit$diagnostics[c("converged", "termination", "iterations", "inner.converged")]
rbind(Theta = dim(fit$Theta), B = dim(fit$B), E.Z = dim(fit$E.Z))
round(fit$Theta[, c(3, 8, 10), drop = FALSE], 3)
```

The retained example completes but reports `loss_increase`; all 25 candidate
fits in the earlier checked run had that status. The historical procedure
retains the increasing trial and can select it. **Successful execution is not
optimizer convergence.** The output above is evaluated again when this article
is built; no stopping rule is changed to make the example look successful.

## Reproduce the historical inference outputs

These are the original example's entrywise and predictor-level outputs. We
retain them as a compatibility demonstration, with the fit's warning still
visible. They are not evidence of inferential calibration or recommendations
to interpret p-values from an unconverged analysis.

```{r historical-inference}
pval.entry <- entry.pvalue(X, Y, fit$Theta, fit$B, fit$E.Z, precision)
pval.predictor <- pleio.pvalue(X, Y, fit$Theta, fit$B, fit$E.Z, precision)
dim(pval.entry)        # one value for each outcome–predictor pair
length(pval.predictor) # one combined value for each predictor
signif(pval.predictor, 3)
```

The predictor combination retains its historical **two-sided Cauchy tail**.
Use the `r site_link("methods and status article", "methods-status.html")` for its limitations
and the optional-K issue. All introductory examples use `K = NULL`.

## Read a coefficient on the right scale

Default fitting centers X and Y and divides each column by its sample standard
deviation. Thus `Theta[r, j]` is the fitted outcome-r change in standard-deviation
units per one-standard-deviation change in predictor j, holding the other
observed predictors fixed on the model scale. This is an association, not an
automatic causal effect or an unbiased estimate.

$$
\Theta_{\rm raw}=\operatorname{diag}(s_Y)\,\Theta\,
\operatorname{diag}(s_X)^{-1},\qquad
\alpha=\bar Y-\Theta_{\rm raw}\bar X.
$$

Multiply each outcome row by its original standard deviation, then divide each
predictor column by its original standard deviation:

```{r raw-scale}
sx <- apply(X, 2, sd)
sy <- apply(Y, 2, sd)
Theta.raw <- sweep(sweep(fit$Theta, 1, sy, "*"), 2, sx, "/")
intercept <- colMeans(Y) - drop(Theta.raw %*% colMeans(X))
Y.predictor <- sweep(X %*% t(Theta.raw), 2, intercept, "+")
round(head(Y.predictor, 3), 2)
```

This is the **observed-predictor component**, not the full training fit. The
latent training contribution is `fit$E.Z %*% t(fit$B)` on the standardized
outcome scale. Its scores use observed outcomes and cannot simply be carried
over to new participants. There is no general `predict()` method. Retain training
centers/scales and the predictor order; remove missing or nonfinite values and
handle constant columns before standardizing.

## Scripts, help and a separate baseline

The shorter `r site_link("quickstart script", "../downloads/quickstart.R")` uses a four-cell
remMap grid and one finite-budget fit, then checks shapes and scale conversion:

```{r installed-example, eval=FALSE}
source(system.file("examples", "quickstart.R", package = "drfarm"))
```

The `r site_link("original reproduction runner", "../downloads/reproduce.R")` records the full
example's data, fitted objects, warnings, timings and environment. From a source
checkout, using a new output directory:

```sh
Rscript --vanilla tools/reproduce.R --mode full --output original-demo \
  --label original-bundled-example
```

The `r site_link("Gaussian ECM example", "../downloads/gaussian-ecm-reference.R")` is a separate
deterministic optimization fixture with explicit starts and no automatic
standardization. Its recorded small fit converged after 65 cycles. It removes
inner debiasing and supplies no inference; it does not replace the example above.

```{r help, eval=FALSE}
source(system.file("examples", "gaussian-ecm-reference.R", package = "drfarm"))
help(package = "drfarm")
?DrFARM.whole
citation("drfarm")
```

Package authors are Lap Sum Chan, Gen Li and Peter X.K. Song. The installed
software citation records the version; GPL ≥ 3 and third-party notices are
preserved. The `r site_link("function reference", "../reference/index.html")` and
`r site_link("citation page", "../authors.html#citation")` provide the full details.
