---
title: "Getting started with drfarm"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with drfarm}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## What is estimated?

For continuous responses, DrFARM represents the data as

$$Y = X\Theta^T + ZB^T + E.$$

`X` has one row per participant and one column per observed predictor; `Y` has
one column per outcome. The fitted `Theta` has **outcome rows and predictor
columns**. A sparse-group penalty encourages sparsity in its entries and
predictor columns. The factor structure describes residual outcome dependence,
not a low-rank constraint on `Theta`. Choosing the number of factors is a
modeling decision. This vignette uses the package's simulated rank-2 example.

## Run the bounded example

Install this maintenance candidate first. The following executes the same
script distributed in `inst/examples/quickstart.R`: a four-candidate remMap
grid, one precision estimate, and one DrFARM fit with a finite iteration budget.

```{r quickstart}
source(system.file("examples", "quickstart.R", package = "drfarm"), local = TRUE)
```

The script prints coefficients, termination diagnostics, the observed-predictor
component on the input scale, and the R session. A returned fit is not itself a
convergence certificate. Inspect `fit$diagnostics`: its convergence flag requires
the historical outer loss-change and inner coefficient-change rules, with no
KKT or global-optimum assertion.
An iteration limit or loss increase is a distinct termination reason. The
historical loss-increase trial-return behavior is retained.

This small grid illustrates the API; it does not replace tuning for a scientific
analysis. `DrFARM.grid()` and `DrFARM.EBIC()` support explicit candidate fitting
and selection, while `DrFARM.whole()` performs its built-in grid search. Preserve
the candidate order, preprocessing, factor initialization options, and selection
criterion when comparing results.

## Keep the scales straight

With the default `standardize = TRUE`, each column is centered and divided by
its sample standard deviation. The fitted coefficients map standardized
predictors to standardized outcomes. If `sx` and `sy` are the original column
standard deviations, then

$$\Theta_{\mathrm{raw}} = \operatorname{diag}(s_y)\,
\Theta\,\operatorname{diag}(s_x)^{-1},\qquad
\alpha = \bar Y-\Theta_{\mathrm{raw}}\bar X.$$

The quickstart verifies that this back-transformation matches transformation of
the standardized predictor component. This algebra does not convert the sparse
estimate into an unbiased one. The stored generating coefficients
`drfarm.dat$Theta.t` have the opposite, predictor-by-outcome orientation; matching
them also requires matching scale.

For `K = NULL`, the training latent contribution is `fit$E.Z %*% t(fit$B)` on the
standardized outcome scale. Those scores use the training outcomes, so they
cannot simply be reused for prediction on new participants. Retain training
centers/scales and participant/predictor/outcome identifiers. The current API
does not supply a general `predict()` method.

## What this example does not establish

The stored simulated data are bundled with the original package. The original
simulation script and seed are not bundled, so this example reproduces an
installation-to-result path rather than the original simulation study.

The original entrywise outer-debiasing and predictor-level p-value functions are
still available. Their use requires the matching scale, precision and variance
assumptions. The historical predictor combination uses a two-sided Cauchy tail;
it is not standard one-sided ACAT. Its inferential interpretation and the
optional-`K` fitting/selection basis are under review. This example therefore
uses `K = NULL` and reports estimates and status. No generalized-response
inference is introduced here.

Use `help(package = "drfarm")` for the API and `citation("drfarm")` for the
package citation. Original authorship and existing GPL notices are preserved.
